Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-10 paper/broom.tex
Related work.
Motivating examples added.
Added bib file.
Updates to abstract and introduction

\newcommand{\absof}[1]{\left\| #1 \right\| }

@@

\newcommand{\angR}{\inang{\rhoalloc\rhobar \,|\, \phi}}

@@

\newcommand{\mang}{\inang{\rhoalloc_m\bar{\rho_m} \,|\, \phi_m}}

@@

@@

\newcommand{\lstjava}{\lstset{ %
language=Java, % choose the language of the code
basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
%escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={letregion, let, in, as, openalloc, open, withroot, not, : , /\\}
}}
\lstnewenvironment{codejava}
    { % \centering
			\lstjava
      \lstset{mathescape=true}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

@@

\begin{abstract}

Distributed execution engines, such as MapReduce and Dryad, represent
a data-parallel computation as a dataflow graph with computational
vertices connected by communication edges. The computation is run by
executing the vertices of this graph on a set of available computers,
while communication edges are materialized as either network
connections, or appropriate inter-process communication channels.
Although MapReduce and Dryad can perform optimal distribution the
computation across available computers, they nevertheless rely on
language runtimes, such as JVM or C\#.Net, to manage an atomic
computation on a given machine.

In this paper, we demonstrate that the general purpose garbage
collection scheme employed by C\#.Net runtime is not very effective
for dataflow programs to the extent that the latency due to GC is
intolerably high in many practical examples. We propose a memory
management scheme that combines user-managed memory regions with
automatic garbage collection, and lets programmers exploit specific
memory behaviours exhibited by dataflow programs. To protect
programmers from the dangers of manual memory management, we
supplement our memory management scheme with a region type system that
statically ensures the memory safety of region-annotated C\# programs,
even in presence of higher-order stateful functions. Further, we
describe a type inference algorithm that infers principal region types
for first-order programs thereby saving programmers the need to write
region annotations on types. Our inference algorithm is also very
effective, albeit incomplete, for higher-order programs. As a
testament to efficacy of our approach, we demonstrate performance
improvements over a range of realistic dataflow programs that were
rewritten to use region-based manual memory management, but are
nevertheless judged to be safe by our type system.

\end{abstract}

@@

% \appendix
% \section{Appendix Title}

@@

\bibliographystyle{abbrvnat}
\bibliography{catalyst}

>>>>

%\newcommand{\absof}[1]{\left\| #1 \right\| }
\newcommand{\absof}[1]{\thesemof{#1}}

@@

\newcommand{\angR}{\inang{\rhoalloc,\rhobar \,|\, \phi}}

@@

\newcommand{\mang}{\inang{\rhoalloc_m,\bar{\rho_m} \,|\, \phi_m}}

@@

\newcommand{\phictxt}{\phi_{cx}}
\newcommand{\phicstr}{\phi_{cs}}
\newcommand{\phisol}{\phi_{sol}}

@@

\newcommand{\lstjava}{\lstset{ %
language=Java, % choose the language of the code
basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
%escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={var, unit, letregion, let, in, as, openalloc, open, withroot, not, : , /\\}
}}
\lstnewenvironment{codejava}
    { % \centering
			\lstjava
      \lstset{mathescape=true}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

@@

\begin{abstract}
\input{abstract}
\end{abstract}

@@

\input{related}
% \appendix
% \section{Appendix Title}

@@

\bibliographystyle{abbrvnat}
\bibliography{broom}

### 2015-11-10 paper/introduction.tex
Updates to abstract and introduction

\begin{itemize}
  \item We present \name, a \csharp-like typed object-oriented language
  that eschews garbage collection in favour of programmer-managed
  memory regions . \name extends its core language, which includes
  \emph{lambdas} (higher-order functions) and \emph{generics}
  (parametric polymorphism), with constructs to create, manage and
  destroy static and dynamic memory regions. Dynamic regions are
  first-class values in \name; they manifest as objects of type
  \C{Region}, and are treated on par with other objects. \name is
  particularly suited for implementing distributed dataflow operators
  (e.g: \c{SELECT} and \c{JOIN}), whose memory behavior exhibits
  certain characteristics that admits efficient and \emph{mostly} safe
  region-based memory management. Experiments demonstrate that
  \naiad~\cite{naiad} dataflow operators using programmer-managed
  memory regions outperform their GC counterparts by a margin of upto
  59\%.

  \item \name is equipped with a region type system that statically
  guarantees safety all memory accesses in a well-typed program,
  subject to some conditions that can be checked efficiently at
  run-time. The overhead of checking these conditions is less than
  \GK{x\%} in our experiments with \naiad workloads.

>>>>

\begin{itemize}
  \item We present \name, a \csharp-like typed object-oriented language
  that eschews garbage collection in favour of programmer-managed
  memory regions . \name extends its core language, which includes
  \emph{lambdas} (higher-order functions) and \emph{generics}
  (parametric polymorphism), with constructs to create, manage and
  destroy static and dynamic memory regions. Dynamic regions are
  first-class values in \name; they manifest as objects of type
  \C{Region}, and are treated on par with other objects. \name is
  particularly suited for dataflow programs, such as~\naiad,
  whose memory behavior exhibits
  certain characteristics that admits efficient region-based memory management.
  Experiments demonstrate that   \naiad~\cite{naiad} dataflow program using programmer-managed
  memory regions outperform their GC counterparts by a margin of upto
  59\%.

  \item \name is equipped with a region type system that statically
  guarantees safety of all memory accesses in a well-typed program,
  subject to some conditions that can be checked efficiently at
  run-time.
  The overhead of checking these conditions is less than
  \GK{x\%} in our experiments with \naiad workloads.
